---
title: "Lab 11 - Cross-Validation"
author: Tejas Dhomne
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package 
```{r}
library(boot)
```

## 1 create a data frame
```{r}
# load the data
housing.df <- read.csv("BostonHousing.csv")
# first six rows 
head(housing.df)
# column names 
colnames(housing.df)
```

## 2 leave-one-out cross-validation 

### 2.1 model 1: model MEDV as a cubic function of DIS 
```{r}
# fit a regression on the training set
glm1 <- glm(MEDV~poly(DIS,3),data=housing.df)
summary(glm1)
# compute the leave-one-out cross-validation predictor error
loocv.err1 <- cv.glm(housing.df,glm1)
# cross-validated MSE 
loocv.err1$delta[1]
```

### 2.2 model 2: add a 4th degree polynomial in NOX as predictors, in addition to predictors in model 1
```{r}
# fit a regression on the training set 
glm2 <- glm(MEDV~poly(DIS,3)+poly(NOX,4),data=housing.df)
summary(glm2)
# compute the leave-one-out cross-validation predictor error
loocv.err2 <- cv.glm(housing.df,glm2)
# cross-validated MSE 
loocv.err2$delta[1]
```

### 2.3 Model 3: add a 5th degree polynomial in CRIM as predictors, in addition to predictors in Model 2
```{r}
# fit a regression on the training set
glm3 <- glm(MEDV~poly(DIS,3)+poly(NOX,4)+poly(CRIM,5),data=housing.df)
summary(glm3)
# compute the leave-one-out cross-validation predictor error
loocv.err3 <- cv.glm(housing.df,glm3)
# cross-validated MSE 
loocv.err3$delta[1]
```

### 2.4 identify the best a model according to the leave-one-out cross-validation method 
```{r}
# which model is best for predicting median home values?  

# Answer - Model_2 is best for predicting median home values.
# Reason - The cross-validated mean squared error value for Model_1 is 76.68154, Model_2 is 66.39615 and Model_3 is Model_3 is 131.7806. Model_2 is the best model out of the three models for predicting median home value because it has the least cross-validated Mean Squared Error Value i.e. 66.39615. 
```

## 3 k-fold cross validation

### 3.1 model 1: model MEDV as a cubic function of DIS 
```{r}
# compute the 5-fold cross-validation predictor error
set.seed(1)
kfcv.err1 <- cv.glm(housing.df,glm1,K=5)
# cross-validated MSE 
kfcv.err1$delta[1]
```

### 3.2 model 2: add a 4th degree polynomial in NOX as predictors, in addition to predictors in model 1 
```{r}
# compute the 5-fold cross-validation predictor error
set.seed(1)
kfcv.err2 <- cv.glm(housing.df,glm2,K=5)
# cross-validated MSE 
kfcv.err2$delta[1]
```

### 3.3 Model 3: add a 5th degree polynomial in CRIM as predictors, in addition to predictors in Model 2
```{r}
# compute the 5-fold cross-validation predictor error
set.seed(1)
kfcv.err3 <- cv.glm(housing.df,glm3,K=5)
# cross-validated MSE 
kfcv.err3$delta[1]
```

### 3.4 identify the best a model according to the 5-fold cross-validation method 
```{r}
# which model is best for predicting median home values?  

# Answer - Model_2 is best for predicting median home values.
# Reason - The cross-validated mean squared error value for Model_1 is 77.48177, Model_2 is 67.086 and Model_3 is Model_3 is 120.7622. Model_2 is the best model out of the three models for predicting median home value because it has the least cross-validated Mean Squared Error Value i.e. 203.9578. Due to the 5th degree polynomial in Model_3 the line tries to overfit the model causing large cross-validated mean squared error.
```