---
title: "Lab-02: Introduction to R programming"
output:
  html_document:
    toc: FALSE
    toc_depth: 2
    toc_float: true
    theme: simplex
    fig_caption: true
    number_sections: true
---
`

# R Syntax
Why do the following code chunks not work? In each case, correct the code and briefly explain the cause of the error.

Note: you can run each code chunk by itself by clicking the "Run" button at the top right of the code chunk, or using the shortcut Ctrl+Shift+Enter. 

## Variables
Correct any syntax errors in the following code chunk.
```{r}
my_variable <- 10
print(my_variable)
# The error means that R could not find the variable mentioned in the error message due to the wrong variable name i.e. variab1e.
```

## Relational operations
Why doesn't R think that `0.3 + 0.6` equals `0.9`?
```{r}
0.3 + 0.6 == 0.9
```
Answer: Binary floating point math is like this. In most programming languages, it is based on the IEEE 754 standard. The crux of the problem is that numbers are represented in this format as a whole number times a power of two; rational numbers whose denominator is not a power of two cannot be exactly represented.

In the following code chunk, write a statement to test if two numbers are nearly equal. Use it to relate `0.3 + 0.6` to `0.9`. Note: a good solution should report either `TRUE` or `FALSE`.
```{r}
isTRUE(all.equal(0.3+0.6, 0.9))
```

## Conditional statement
Correct any syntax errors in the following code chunk.
```{r}
if (1 > FALSE) {
  print("1 > FALSE")
} else {
  print("1 <= FALSE")
}
# Correct Syntax is Given below:
# if (test_expression) {
# statement1
#  } else {
# statement2
# }
# The else part is optional and is only evaluated if test_expression is FALSE. It is important to note that else must be in the same line as the closing braces of the if statement.
```


## Loop
Correct any syntax errors in the following code chunk.
```{r}
for (i in list("a", TRUE, 0)) {
  print(i)
}

# Correct syntax format of a for loop is
# for (val in sequence)
# {
# statement
# }
# Therefore corrected "from" to "in" as per the syntax, Here, list is a vector and i takes on each of its value during the loop. In each iteration, statement is evaluated.
```


# Subsetting
Below each comment in the following code chunk, add code to perform the requested [subsetting](https://adv-r.hadley.nz/subsetting.html) operations.
```{r}
v <- c(3, 1, 8, 45, 2, 4, 0)

# Use positive integers to subset v to the 2nd and 5th elements
v[c(2, 5)]

# Use negative integers to exclude the 2nd and 5th elements of v
v[-c(2, 5)]

# Use a logical vector to select the first and last elements of v
v[c(TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE)]

# Use the following expression to subset v to elements that are greater than 2
v[v > 2]

# Use the head() function to display the first 6 elements of v
head(v)

# Use the head() function to display the first 3 elements of v
head(v,n=3)

```


# Variable Operations
Add to or modify the code in each code chunk to implement the step described above each chunk.

1. Create an atomic vector `y` with the following numeric values: 5, -2, 31, 4, 32, 7, -9.
```{r}
y <-  c(5, -2, 31, 4, 32, 7, -9)
```

2. Calculate the average value of `y`, and assign it to a new variable `y_mean`.
```{r}
y_mean <- mean(y)
y_mean
```

3. Create a new vector `y_error` which gives the difference between each value of `y` and `y_mean`. 
```{r}
y_error <- y - y_mean
y_error
```

4. Using the `length()` function, construct a single expression that confirms two things: (a) `y_error` and `y` both contain the same number of elements, and (b) the mean of `y_error` is equal to zero (up to machine precision). In other words, evaluating this expression should generate a value of `TRUE` if both conditions (a) and (b) are true, but a value of `FALSE` if either condition is false.
```{r}
# (a)
length(y_error) == length(y)

# (b)
mean(y_error) == 0

# (a) and (b)
if (length(y_error) == length(y)&mean(y_error) == 0) {
print('TRUE')
} else {
print('FALSE')
}
```

5. Create a new vector `y_abs_error` which gives the absolute value of element of `y_error`. Calculate and print the mean of the absolute error variable.
```{r}
y_abs_error <- abs(y_error)
print(y_abs_error)
  
```

# Functions
Create a function called `to_zero` that takes a numeric value `x` and subtracts 1 if the input value is greater than 0, adds 1 if the input is less than zero, and returns zero if the input is equal to zero. Follow the comments below.
```{r}
# Define the function to use a default value of x = 2
to_zero <- function(x=2) {
  if (x>0){
    x-1
  } else if(x<0){
    x+1
  } else {
    0
  }
  
}

# Evaluate the function with no argument. Also evaluate the function at the values 1, 0, and -1.
to_zero()
to_zero(1)
to_zero(0)
to_zero(-1)
```


# Data Frames
The following exercises involve reading text data into a data frame. You may wish to refer to the lecture handout or to the [Data import](https://r4ds.had.co.nz/data-import.html) chapter of the textbook.

## Read a delimited (csv) text file into a data frame
Use the `read_csv()` function from the `tidyverse` package to load the CSV file `20160819_OlympicSportsByCountries.csv` available at the web address below. For R to recognize the `read_csv` command, you must first load the `tidyverse` package, e.g. by using the `require` command. 

**HINT:** After the code chunk below has been correctly modified, the data frame `trends_csv` should have three variables (Country, Search Interest, and Sport) and 1,408 rows. You will need to modify the arguments of the `read_csv` command below to read in the data file without error. You may want to open the raw data file in a text editor to gain insight into what may be causing the error. Enter `?read_csv` in the R Console to learn more about the `read_csv` command. 

Original code:
```{r, eval = TRUE}
library(readr)
require(tidyverse)

url <- "https://raw.githubusercontent.com/googletrends/data/master/20160819_OlympicSportsByCountries.csv"
trends_csv <- read_csv(url, skip = 2, comment = "=", n_max = 1408, guess_max = 1408, col_types = cols("Search Interest" = col_double()))
all.equal(names(trends_csv), c("Country", "Search Interest", "Sport"))
nrow(trends_csv)==1408
```


## Preview data in data frame
Print the first three observations (i.e. first three rows) in the data frame `trends_csv` created above. Do this using the `head` function, and also do it using index notation as described in the section on Selecting Observations [here](https://www.statmethods.net/management/subset.html). 
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```
```{r}

head(trends_csv,3)
trends_csv[1:3,]
```

## Subsetting a data frame
Using the data frame `trends_csv` created above, select and list all observations corresponding to the country of China. Which sport is most searched for in China?
```{r}
filter(trends_csv, Country == "China")

trends_csv %>%
  filter(Country == 'China') %>%
  slice(which.max(`Search Interest`))
# Swimming is most searched in China where Search Interest = 15
```

## Going further
**Advanced (and optional):** Write a function called `top_sport()` that takes the name of a country as a single (string) argument and prints the name of the most searched for sport in that country based on data in the data frame `trends_csv` created above. E.g., `top_sport("United States")` should print the name of the most searched for sport in the U.S. If you figure this out, feel free to email the professor your proposed solution!
```{r}
top_sport <- function(x) {
 trends_csv %>%
  filter(Country == x) %>%
  slice(which.max(`Search Interest`))
}
  
top_sport("United States")
top_sport("China")
top_sport("Russia")
```


