---
title: "Lab 16 - Neural Nets (Solution)"
author: Tejas Dhomne
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
knitr::knit_meta(clean=T)
```

## 0 load the packages 
```{r}
library(neuralnet)
library(caret)
library(gains)

```

## 1 create a data frame
```{r}
# load the data 
df <-read.csv("EastWestAirlinesNN.csv")
# first six rows 
head(df)
# column names
names(df)
```

## 2 create two outcome dummies to represent output nodes  
```{r}
# create a variable named purchase
df$purchased <- ifelse(df$Phone_sale == "1",TRUE, FALSE)
# create a variable named not_purchase
df$not_purchase <- ifelse(df$Phone_sale == "0",TRUE,FALSE)
df
```

## 3 remove unnecessary variables
```{r}
# remove ï..ID column and update df 
df = subset(df[,-1])
# column names 
names(df)
```

## 4 remove rows with missing values
```{r}
df <- na.omit(df)

```

## 5 data partition 
```{r}
# set the seed 
set.seed(1)
# row indexes of the training set
train.index <- sample(c(1:dim(df)[1]), 0.6*dim(df)[1])  
head(train.index)
# training set
train.df <- df[train.index, ]
head(train.df)

# test set
test.df <- df[-train.index, ]
head(test.df)
```

## 6 normalize predictors to a scale of [0, 1] 
```{r}
# estimate the transformation 
library(caret)
norm.values <- preProcess(train.df[,c(1:14)], method="range")
# normalize the numerical predictors in the training set 
train.norm.df <- predict(norm.values, train.df[,c(1:14)])
head(train.norm.df)
# normalize the numerical predictors in the test set 
test.norm.df <- predict(norm.values, test.df[,c(1:14)])
head(test.norm.df)
```

## 7 fit a neural network on the training set 
```{r}
# set the random seed 
set.seed(1)
# fit a neural network that has a single hidden layer with 5 nodes 
nn <- neuralnet(train.df$not_purchase + train.df$purchase ~ Topflight + Balance + Qual_miles + cc1_miles. + cc2_miles. + cc3_miles. + Bonus_miles + Bonus_trans + Flight_miles_12mo + Flight_trans_12 + Online_12 + Email +
Club_member +  Any_cc_miles_12mo, data = train.norm.df, linear.output = FALSE, hidden = 5)
# plot the network
plot(nn, rep = "best")
# display weights
nn$weights
```

## 8 generate predicted probabilities for records in the test set  
```{r}
# make predictions using compute() 
nn.pred <- compute(nn, test.norm.df)
# predicted probabilities of making purchases
nn.pred.prob <- nn.pred$net.result[,2]/(nn.pred$net.result[,1]+nn.pred$net.result[,2])
```

## 9 generate predicted classes for records in the test set 
```{r}
nn.pred.class <- ifelse(nn.pred.prob>0.5,1,0)
nn.pred.class
```

## 10 create a confusion matrix for records in the test set
```{r}
confusionMatrix(as.factor(nn.pred.class), as.factor(test.df$Phone_sale),positive="1")

```

## 11 create a gain table 
```{r}
# gain table 
gain <- gains(as.numeric(as.character(test.df$Phone_sale)), nn.pred.prob, groups = 10)
# cumulative percentage of customers who make purchases 
gain$cume.pct.of.total
# cumulative number of customers 
gain$cume.obs
```

## 12 plot a lift chart
```{r}
# plot the cumulative number of customers who make purchases against the cumulative number of customers
plot(c(0,gain$cume.pct.of.total*sum(test.df$Phone_sale))~c(0,gain$cume.obs), 
     xlab="cumulative number of customers", ylab="cumulative number of customers who make purchases", type="l")
# add a baseline curve 
lines(c(0,sum(test.df$Phone_sale))~c(0, dim(test.df)[1]))
```